---
title: "Untitled"
author: "utomoreza"
date: "07/02/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(ggthemes)
library(plotly)
library(scales)
```

```{r}
# define a function to add 'years' at the end of a string
add_year <- function(x) {return(paste(x, 'years'))}

# define a function to convert factor to numeric
as.numeric.factor <- function(x) {as.numeric(levels(x))[x]}
```

```{r}
countries1 <- as.data.frame(read_tsv("./LifeExpectancy/countries.tsv", na = ""))

countries2 <- as.data.frame(read_tsv("./LifeExpectancy/countriesGoogle.tsv", na = "")) %>% 
        select(country, latitude, longitude)

countries <- left_join(countries1, countries2, by = c("Country Code" = "country"))

for (i in 1:length(countries$latitude)) {
        if (is.na(countries$latitude[i])) {
                countries$latitude[i] <- replace_na(countries$latitude[i], 
                                                    countries$`Capital Latitude`[i])
                countries$longitude[i] <- replace_na(countries$longitude[i], 
                                                    countries$`Capital Longitude`[i])
        }        
}
countries_ <- countries %>% 
        rename(Country = `Country Name`,
               Capital = `Capital Name`,
               Continent = `Continent Name`,
               Latitude = latitude,
               Longitude = longitude) %>% 
        select(Country, Capital, Continent, Latitude, Longitude) %>% 
        mutate(Continent = as.factor(Continent))
countries_$Country[52] <- "Cote d'Ivoire"
countries_$Capital[125] <- "Macau"
countries_$Country[174] <- "Democratic Republic of Congo"
countries_$Country[217] <- "Gambia"
```

# First Dataset

```{r}
lifeExp <- as.data.frame(read_csv("./LifeExpectancy/life-expectancy.csv"))
lifeExp
```

```{r}
lifeExp_countries <- lifeExp %>% 
     filter((!is.na(Code) | Entity == "Saint Barthlemy") & Entity != "World") %>% 
     rename(Country = Entity, LE_age = `Life expectancy (years)`) %>% 
     mutate(Country = as.factor(Country), Code = as.factor(Code))
lifeExp_regions <- lifeExp %>% 
     filter((is.na(Code) | Entity == "World") & Entity != "Saint Barthlemy") %>% 
     select(Entity, Year, `Life expectancy (years)`) %>% 
     rename(Region = Entity, LE_age = `Life expectancy (years)`) %>% 
     mutate(Region = as.factor(Region))
lifeExp_countries
lifeExp_regions
```

2. Menampilkan grafik perbandingan pertumbuhan life expectancy negara-negara tertentu pada rentang waktu tertentu

```{r}
# Select only 5 countries in maximum
country1 <- "Indonesia"
country2 <- "Malaysia"
country3 <- "Singapore"
country4 <- "Myanmar"
country5 <- "Thailand"

# subset by countries chosen
lifeExp_countries_chosen <- lifeExp_countries %>% 
     filter(Country == country1 | 
                 Country == country2 | 
                 Country == country3 | 
                 Country == country4 | 
                 Country == country5)

p1a <- ggplot(lifeExp_countries_chosen, aes(x = Year,
                              y = LE_age,
                              text = paste0(Country, "<br>",
                                            "Year: ", Year, "<br>",
                                            "Life expectancy: ", round(LE_age, 1)),
                              group = 1)) +
     geom_line(aes(color = Country)) +
     labs(title = "Life expectancy",
          caption = "Source: ourworldindata.org") +
     scale_y_continuous(breaks = seq(from = 0, to = 90, by = 10), 
                        labels = add_year) +
     theme_minimal() +
     theme(axis.title.x = element_blank(),
           axis.title.y = element_blank(),
           plot.title = element_text(face = "bold", size = 18),
           plot.caption = element_text(size = 7),
           legend.title = element_blank())

plotly1a <- ggplotly(p1a, tooltip = "text")
plotly1a
```

3. Menampilkan grafik perbandingan pertumbuhan life expectancy semua benua pada rentang waktu tertentu

```{r}
# lifeExp_regions_wide <- lifeExp_regions %>% 
#         pivot_wider(names_from = "Region", values_from = "LE_age")



p1b <- ggplot(lifeExp_regions, aes(x = Year,
                              y = LE_age,
                              text = paste0(Region, "<br>",
                                            "Year: ", Year, "<br>",
                                            "Life expectancy: ", round(LE_age, 1)),
                              group = 1)) +
     geom_line(aes(color = Region)) +
     labs(title = "Life expectancy",
          caption = "Source: ourworldindata.org") +
     scale_y_continuous(breaks = seq(from = 0, to = 90, by = 10), 
                        labels = add_year) +
     theme_minimal() +
     theme(axis.title.x = element_blank(),
           axis.title.y = element_blank(),
           plot.title = element_text(face = "bold", size = 18),
           plot.caption = element_text(size = 7),
           legend.title = element_blank(),
           legend.position = "top")

plotly1b <- ggplotly(p1b, tooltip = "text")
plotly1b
```

# Second Dataset

```{r}
MenVSWomen <- as.data.frame(read_csv("./LifeExpectancy/life-expectancy-of-women-vs-life-expectancy-of-women.csv"))
MenVSWomen
```

```{r}
MenVSWomen_countries <- MenVSWomen %>% 
     filter((!is.na(Code) | Entity == "Saint Barthlemy") & Entity != "World") %>% 
     rename(Country = Entity, 
            LE_men_age = `Life expectancy of men (years)`,
            LE_women_age = `Life expectancy of women (years)`,
            Total_pop = `Total population (Gapminder)`) %>% 
     mutate(Country = as.factor(Country),
            Code = as.factor(Code))
MenVSWomen_regions <- MenVSWomen %>% 
     filter((is.na(Code) | Entity == "World") & Entity != "Saint Barthlemy") %>%
     select(Entity, 
            Year, 
            `Life expectancy of men (years)`, 
            `Life expectancy of women (years)`, 
            `Total population (Gapminder)`) %>% 
     rename(Region = Entity, 
            LE_men_age = `Life expectancy of men (years)`,
            LE_women_age = `Life expectancy of women (years)`,
            Total_pop = `Total population (Gapminder)`) %>% 
     mutate(Region = as.factor(Region))
MenVSWomen_countries
MenVSWomen_regions
```

4. Menampilkan grafik perbandingan life expectancy laki-laki dan perempuan di negara-negara dunia

```{r}
# Choose year to plot
sliceYear <- 2000

MenVSWomen_countries_yearChosen <- MenVSWomen_countries %>% 
     filter(Year == sliceYear)

p2a <- ggplot(MenVSWomen_countries_yearChosen, aes(x = LE_men_age,
                                 y = LE_women_age,
                                 size = Total_pop,
                                 text = paste0(Country, "<br>",
                                               "Men: ", round(LE_men_age,2), "<br>",
                                               "Women: ", round(LE_women_age,2), "<br>",
                                               "Pop: ", format(Total_pop, big.mark = ",")))) +
     geom_jitter(aes(color = Country), alpha = 0.8) +
     labs(title = paste0("Men's vs women's life expectancy in ", sliceYear),
          x = "Life expectancy of men",
          y = "Life expectancy of women") +
     theme_bw() +
     theme(legend.position = "none",
           plot.title = element_text(face = "bold", size = 13))

ggplotly(p2a, tooltip = "text")
```

```{r}
MenVSWomen_countries[MenVSWomen_countries$Year == 2015,]
```

```{r}
MenVSWomen_regions_yearChosen <- MenVSWomen_regions %>% 
     filter(Year == sliceYear)

p2b <- ggplot(MenVSWomen_regions_yearChosen, aes(x = LE_men_age,
                                                 y = LE_women_age,
                                                 size = Total_pop,
                                                 text = paste0(Region, "<br>",
                                               "Men: ", round(LE_men_age,2), "<br>",
                                               "Women: ", round(LE_women_age,2), "<br>",
                                               "Pop: ", format(Total_pop, big.mark = ",")))) +
     geom_jitter(aes(color = Region), alpha = 0.8) +
     scale_size_continuous(labels = comma) +
     labs(title = paste0("Men's vs women's life expectancy in ", sliceYear),
          x = "Life expectancy of men",
          y = "Life expectancy of women") +
     theme_bw() +
     theme(legend.position = "none",
           plot.title = element_text(face = "bold", size = 13))

ggplotly(p2b, tooltip = "text")
```

```{r}
MenVSWomen_countries_long <- MenVSWomen_countries %>% 
     pivot_longer(cols = c("LE_men_age","LE_women_age"),
                  names_to = "Gender", values_to = "LE_age") %>% 
        as.data.frame()

MenVSWomen_countries_long$Gender <- as.factor(sapply(as.character(MenVSWomen_countries_long$Gender), switch,
                                                     "LE_men_age" = "Men",
                                                     "LE_women_age" = "Women"))

# Choose a country
countryChosen <- "United Kingdom"

dfLong_counChosen <- MenVSWomen_countries_long %>% 
     filter(Country == countryChosen) %>% 
     drop_na() %>% 
     as.data.frame()

p3a <- ggplot(dfLong_counChosen, aes(x = Year,
                                     y = ifelse(Gender == "Men",
                                                LE_age, 
                                                -LE_age),
                                     fill = Gender,
                                     text = paste0(Year, "<br>",
                                            "Life exp: ", round(LE_age,2),
                                            " years"))) +
        geom_col() +
        coord_flip() + 
        # scale_fill_manual(labels = c("Men","Women"), values = c("blue","red")) +
        scale_fill_brewer(palette = "Set1") +
        scale_y_continuous(labels = abs,
                           limits = max(dfLong_counChosen$LE_age)*c(-1,1)) +
        labs(title = paste0("Life expectancy by gender in ", countryChosen),
             y = "Life expectancy",
             x = NULL) +
        theme_calc() +
        theme(plot.title = element_text(face = "bold", size = 12))

ggplotly(p3a, tooltip = "text") 
```

```{r}
MenVSWomen_regions_long <- MenVSWomen_regions %>% 
     pivot_longer(cols = c("LE_men_age","LE_women_age"),
                  names_to = "Gender", values_to = "LE_age") %>% 
        as.data.frame()

MenVSWomen_regions_long$Gender <- as.factor(sapply(
        as.character(MenVSWomen_regions_long$Gender), switch,
        "LE_men_age" = "Men",
        "LE_women_age" = "Women"))

# Choose a region
regionChosen <- "Africa"

dfLong_regChosen <- MenVSWomen_regions_long %>% 
     filter(Region == regionChosen) %>% 
     drop_na() %>% 
     as.data.frame()

p3b <- ggplot(dfLong_regChosen, aes(x = Year,
                                     y = ifelse(Gender == "Men",
                                                LE_age, 
                                                -LE_age),
                                     fill = Gender,
                                     text = paste0(Year, "<br>",
                                            "Life exp: ", round(LE_age,2),
                                            " years"))) +
        geom_col() +
        coord_flip() + 
        # scale_fill_manual(labels = c("Men","Women"), values = c("blue","red")) +
        scale_fill_brewer(palette = "Set1") +
        scale_y_continuous(labels = abs,
                           limits = max(dfLong_regChosen$LE_age)*c(-1,1)) +
        labs(title = paste0("Life expectancy by gender in ", regionChosen),
             y = "Life expectancy",
             x = NULL) +
        theme_calc() +
        theme(plot.title = element_text(face = "bold", size = 12))

ggplotly(p3b, tooltip = "text") 
```

```{r}
df_pop <- MenVSWomen_regions %>% 
        filter(Region != "World" & Region != "Saint Barthlemy") %>% 
        select(Region, Year, Total_pop) %>% 
        drop_na() %>% 
        mutate(Year = as.factor(Year))

Total <- df_pop %>% 
        group_by(Year) %>% 
        summarise(sum_Year = sum(Total_pop)) %>% 
        ungroup()

dfpopPercent <- left_join(df_pop, Total, by = c("Year" = "Year")) %>% 
        mutate(Ratio = Total_pop/sum_Year*100,
               Year = as.numeric.factor(Year)) %>% 
        select(Region, Year, Total_pop, Ratio)
```

```{r}
options(scipen = 999)
p4 <- ggplot(dfpopPercent, aes(x = Year,
                               y = Total_pop,
                               text = paste0(Region, "<br>",
                                             "Year: ", Year, "<br>",
                                             "Pops: ", format(Total_pop, big.mark = ",")),
                               group = Region)) +
        geom_area(aes(fill = Region),
                  position = "fill",
                  color = "white",
                  size = 0.2) +
        labs(title = "World Population - Normalized",
             x = "Year",
             y = NULL) +
        scale_x_continuous(labels = comma) +
        scale_fill_brewer(palette = "Dark2") +
        scale_y_continuous(labels = percent) +
        theme_calc() +
        theme(plot.title = element_text(face = "bold", size = 15))

ggplotly(p4, tooltip = "text")

```

# Third Dataset

```{r}
LEvsGDP <- as.data.frame(read_csv("./LifeExpectancy/life-expectancy-vs-gdp-per-capita.csv"))
LEvsGDP
```

```{r}
LEvsGDP_countries <- LEvsGDP %>% 
     filter(!is.na(Code) & Entity != "World") %>% 
     rename(Country = Entity, 
            LE_age = `Life expectancy at birth (years)`,
            GDP = `GDP per capita ($)`,
            Total_pop = `Population by country`) %>% 
     mutate(Country = as.factor(Country),
            Code = as.factor(Code))
LEvsGDP_regions <- LEvsGDP %>% 
     filter(is.na(Code) | Entity == "World") %>% 
     select(Entity, 
            Year, 
            `Life expectancy at birth (years)`,
            `GDP per capita ($)`,
            `Population by country`) %>% 
     rename(Region = Entity, 
            LE_age = `Life expectancy at birth (years)`,
            GDP = `GDP per capita ($)`,
            Total_pop = `Population by country`) %>% 
     mutate(Region = as.factor(Region))
LEvsGDP_countries
LEvsGDP_regions
```

```{r}
LEvsGDP_countries_ <- LEvsGDP_countries %>% 
        mutate(Country = as.character.factor(Country))
LEvsGDP_countries_$Country <- str_replace_all(LEvsGDP_countries_$Country, 
                                              "Brunei", 
                                              "Brunei Darussalam")
LEvsGDP_countries_$Country <- str_replace_all(LEvsGDP_countries_$Country, 
                                              "Macao", 
                                              "Macau")
LEvsGDP_countries_$Country <- str_replace_all(LEvsGDP_countries_$Country, 
                                              "^Congo$", 
                                              "Democratic Republic of Congo")
LEvsGDP_countries_$Country <- str_replace_all(LEvsGDP_countries_$Country, 
                                              "^Micronesia \\(country\\)$", 
                                              "Federated States of Micronesia")
LEvsGDP_countries_$Country <- str_replace_all(LEvsGDP_countries_$Country, 
                                              "Timor", 
                                              "Timor-Leste")
LEvsGDP_countries_$Country <- str_replace_all(LEvsGDP_countries_$Country, 
                                              "^Polynesia$", 
                                              "French Polynesia")

LEvsGDP_countrFull <- left_join(LEvsGDP_countries_,
                                countries_,
                                by = c("Country" = "Country")) %>% 
        mutate(Country = as.factor(Country)) %>% 
        filter(Country != "Channel Islands" & 
                       Country != "French Guiana" &
                       Country != "Macacu" &
                       Country != "Melanesia" &
                       Country != "Reunion" &
                       Country != "Curacao" &
                       Country != "Martinique" &
                       Country != "Guadeloupe" &
                       Country != "Mayotte" &
                       Country != "United States Virgin Islands")

LEvsGDP_countrFull[9980:10073,"Capital"] <- "Prague"
LEvsGDP_countrFull[9980:10073,"Continent"] <- "Europe"
LEvsGDP_countrFull[9980:10073,"Latitude"] <- 49.81749
LEvsGDP_countrFull[9980:10073,"Longitude"] <- 15.47296

LEvsGDP_countrFull[44622:44841,"Capital"] <- "Vatican"
LEvsGDP_countrFull[44622:44841,"Continent"] <- "Europe"
LEvsGDP_countrFull[44622:44841,"Latitude"] <- 41.87194
LEvsGDP_countrFull[44622:44841,"Longitude"] <- 12.56738

LEvsGDP_countrFull
```

```{r}
# Choose year to plot
sliceYear <- 2010

DF_LEvsGDPCountries <- LEvsGDP_countrFull %>% 
     filter(Year == sliceYear)

p5 <- ggplot(DF_LEvsGDPCountries, aes(x = LE_age,
                                      y = GDP,
                                      size = Total_pop,
                                      text = paste0(Country, "<br>",
                                                    "Life exp: ", round(LE_age,2), "<br>",
                                                    "GDP: $", format(GDP, big.mark = ","), "<br>",
                                                    "Pops: ", format(Total_pop, big.mark = ",")))) +
        geom_jitter(aes(color = Continent), alpha = 0.6) +
        labs(title = paste0("Life expectancy vs GDP per capita in ", sliceYear),
             x = NULL,
             y = NULL) +
        scale_y_continuous(labels = dollar) +
        scale_x_continuous(labels = add_year) +
        scale_color_brewer(palette = "Set1") +
        theme_bw() +
        theme(plot.title = element_text(face = "bold", size = 13))

ggplotly(p5, tooltip = "text")
```

# Fourth Dataset

```{r}
LEvsHealth <- as.data.frame(read_csv("./LifeExpectancy/life-expectancy-vs-healthcare-expenditure.csv", na = ""))
LEvsHealth <- LEvsHealth %>% 
        rename(Country = Entity,
               LE_age = `Life expectancy at birth (years)`,
               HealthExp = `Healthcare Expenditure per capita (int.-$) (constant 2011 international $)`,
               Total_pop = X6) %>% 
        filter(Year >= 1995 & !is.na(Code)) %>% 
        filter(Country != "Channel Islands" & 
                       Country != "French Guiana" &
                       Country != "Macacu" &
                       Country != "Melanesia" &
                       Country != "Reunion" &
                       Country != "Curacao" &
                       Country != "Martinique" &
                       Country != "Guadeloupe" &
                       Country != "Mayotte" &
                       Country != "United States Virgin Islands")
```

```{r}
LEvsHealth$Country <- str_replace_all(LEvsHealth$Country, 
                                      "Brunei", 
                                      "Brunei Darussalam")
LEvsHealth$Country <- str_replace_all(LEvsHealth$Country, 
                                      "Macao", 
                                      "Macau")
LEvsHealth$Country <- str_replace_all(LEvsHealth$Country, 
                                      "^Congo$", 
                                      "Democratic Republic of Congo")
LEvsHealth$Country <- str_replace_all(LEvsHealth$Country, 
                                      "^Micronesia \\(country\\)$", 
                                      "Federated States of Micronesia")
LEvsHealth$Country <- str_replace_all(LEvsHealth$Country, 
                                      "Timor", 
                                      "Timor-Leste")
LEvsHealth$Country <- str_replace_all(LEvsHealth$Country, 
                                      "^Polynesia$", 
                                      "French Polynesia")

LEvsHealth_countrFull <- left_join(LEvsHealth,
                                countries_,
                                by = c("Country" = "Country")) %>% 
        filter(!is.na(Continent) | Country == "World") %>% 
        mutate(Continent = as.character.factor(Continent))

idx <- as.numeric(rownames(LEvsHealth_countrFull[LEvsHealth_countrFull$Country == "World",]))
LEvsHealth_countrFull[idx,"Continent"] <- "World"

LEvsHealth_countrFull$Continent <- as.factor(LEvsHealth_countrFull$Continent)
```

```{r}
# Choose year to plot
sliceYear <- 2013

DF_LEvsHealthCountries <- LEvsHealth_countrFull %>% 
     filter(Year == sliceYear)

p6 <- ggplot(DF_LEvsHealthCountries, aes(x = LE_age,
                                      y = HealthExp,
                                      size = Total_pop,
                                      text = paste0(Country, "<br>",
                                                    "Life exp: ", round(LE_age,2), 
                                                    "<br>",
                                                    "Healthcare Exp: $", 
                                                    format(
                                                            round(
                                                                    HealthExp,2), 
                                                            big.mark = ","), 
                                                    "<br>",
                                                    "Pops: ", format(Total_pop, big.mark = ",")))) +
        geom_jitter(aes(color = Continent), alpha = 0.6) +
        labs(title = paste0("Life expectancy vs healthcare expenditure in ", sliceYear),
             x = NULL,
             y = NULL) +
        scale_y_continuous(labels = dollar) +
        scale_x_continuous(labels = add_year) +
        scale_color_brewer(palette = "Set1") +
        theme_bw() +
        theme(plot.title = element_text(face = "bold", size = 13))

ggplotly(p6, tooltip = "text")
```

# Rest

```{r}
medianAge <- as.data.frame(read_csv("./LifeExpectancy/median-age.csv"))
medianAge
```

```{r}
medianAge_countries <- medianAge %>% 
     filter(!is.na(Code) & Entity != "World") %>% 
     rename(Country = Entity, 
            Median_Age = `UN Population Division (Median Age) (2017) (years)`) %>% mutate(Country = as.factor(Country),
            Code = as.factor(Code))
medianAge_regions <- medianAge %>% 
     filter(is.na(Code) | Entity == "World") %>% 
     select(Entity, 
            Year, 
            `UN Population Division (Median Age) (2017) (years)`) %>% 
     rename(Region = Entity, 
            Median_Age = `UN Population Division (Median Age) (2017) (years)`) %>% mutate(Region = as.factor(Region))
medianAge_countries
medianAge_regions
```

5. Menampilkan grafik perbandingan median age antara negara-negara tertentu pada rentang waktu tertentu

```{r}
medianAge_countries_ <- medianAge_countries %>% 
        filter(Country == "China" |
                       Country == "India" |
                       Country == "United States")

ggplot(medianAge_countries_, aes(x = Country,
                                 y = Median_Age)) +
        geom_boxplot() +
        geom_
```

