---
title: "Untitled"
author: "utomoreza"
date: "07/02/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(ggthemes)
library(plotly)
library(scales)
```

# First Dataset

```{r}
lifeExp <- as.data.frame(read_csv("./LifeExpectancy/life-expectancy.csv"))
lifeExp
```

```{r}
lifeExp_countries <- lifeExp %>% 
     filter(!is.na(Code) & Entity != "World") %>% 
     rename(Country = Entity, LE_age = `Life expectancy (years)`) %>% 
     mutate(Country = as.factor(Country), Code = as.factor(Code))
lifeExp_regions <- lifeExp %>% 
     filter(is.na(Code) | Entity == "World") %>% 
     select(Entity, Year, `Life expectancy (years)`) %>% 
     rename(Region = Entity, LE_age = `Life expectancy (years)`) %>% 
     mutate(Region = as.factor(Region))
lifeExp_countries
lifeExp_regions
```

2. Menampilkan grafik perbandingan pertumbuhan life expectancy negara-negara tertentu pada rentang waktu tertentu

```{r}
# Select only 5 countries in maximum
country1 <- "Indonesia"
country2 <- "Malaysia"
country3 <- "Singapore"
country4 <- "Myanmar"
country5 <- "Thailand"

# subset by countries chosen
lifeExp_countries_chosen <- lifeExp_countries %>% 
     filter(Country == country1 | 
                 Country == country2 | 
                 Country == country3 | 
                 Country == country4 | 
                 Country == country5)

country <- paste0(lifeExp_countries_chosen$Country, "<br>",
               "Year: ", lifeExp_countries_chosen$Year, "<br>",
               "Life expectancy: ", round(lifeExp_countries_chosen$LE_age),1)

p1a <- ggplot(lifeExp_countries_chosen, aes(x = Year,
                              y = LE_age,
                              text = paste0(Country, "<br>",
                                            "Year: ", Year, "<br>",
                                            "Life expectancy: ", round(LE_age, 1)),
                              group = 1)) +
     geom_line(aes(color = Country)) +
     labs(title = "Life expectancy",
          caption = "Source: ourworldindata.org") +
     scale_y_continuous(breaks = seq(from = 0, to = 90, by = 10), 
                        labels = c('0 year', 
                                   '10 years', 
                                   '20 years',
                                   '30 years', 
                                   '40 years', 
                                   '50 years', 
                                   '60 years', 
                                   '70 years', 
                                   '80 years', 
                                   '90 years')) +
     theme_minimal() +
     theme(axis.title.x = element_blank(),
           axis.title.y = element_blank(),
           plot.title = element_text(face = "bold", size = 18),
           plot.caption = element_text(size = 7),
           legend.title = element_blank())

plotly1a <- ggplotly(p1a, tooltip = "text")
plotly1a
```

3. Menampilkan grafik perbandingan pertumbuhan life expectancy semua benua pada rentang waktu tertentu

```{r}
p1b <- ggplot(lifeExp_regions, aes(x = Year,
                              y = LE_age,
                              text = paste0(Region, "<br>",
                                            "Year: ", Year, "<br>",
                                            "Life expectancy: ", round(LE_age, 1)),
                              group = 1)) +
     geom_line(aes(color = Region)) +
     labs(title = "Life expectancy",
          caption = "Source: ourworldindata.org") +
     scale_y_continuous(breaks = seq(from = 0, to = 90, by = 10), 
                        labels = c('0 year', 
                                   '10 years', 
                                   '20 years',
                                   '30 years', 
                                   '40 years', 
                                   '50 years', 
                                   '60 years', 
                                   '70 years', 
                                   '80 years', 
                                   '90 years')) +
     theme_minimal() +
     theme(axis.title.x = element_blank(),
           axis.title.y = element_blank(),
           plot.title = element_text(face = "bold", size = 18),
           plot.caption = element_text(size = 7),
           legend.title = element_blank(),
           legend.position = "top")

plotly1b <- ggplotly(p1b, tooltip = "text")
plotly1b
```

# Second Dataset

```{r}
MenVSWomen <- as.data.frame(read_csv("./LifeExpectancy/life-expectancy-of-women-vs-life-expectancy-of-women.csv"))
MenVSWomen
```

```{r}
MenVSWomen_countries <- MenVSWomen %>% 
     filter(!is.na(Code) & Entity != "World") %>% 
     rename(Country = Entity, 
            LE_men_age = `Life expectancy of men (years)`,
            LE_women_age = `Life expectancy of women (years)`,
            Total_pop = `Total population (Gapminder)`) %>% 
     mutate(Country = as.factor(Country),
            Code = as.factor(Code))
MenVSWomen_regions <- MenVSWomen %>% 
     filter(is.na(Code) | Entity == "World") %>% 
     select(Entity, 
            Year, 
            `Life expectancy of men (years)`, 
            `Life expectancy of women (years)`, 
            `Total population (Gapminder)`) %>% 
     rename(Region = Entity, 
            LE_men_age = `Life expectancy of men (years)`,
            LE_women_age = `Life expectancy of women (years)`,
            Total_pop = `Total population (Gapminder)`) %>% 
     mutate(Region = as.factor(Region))
MenVSWomen_countries
MenVSWomen_regions
```

4. Menampilkan grafik perbandingan life expectancy laki-laki dan perempuan di negara-negara dunia

```{r}
# Choose year to plot
sliceYear <- 2000

MenVSWomen_countries_yearChosen <- MenVSWomen_countries %>% 
     filter(Year == sliceYear)

p2a <- ggplot(MenVSWomen_countries_yearChosen, aes(x = LE_men_age,
                                 y = LE_women_age,
                                 size = Total_pop,
                                 text = paste0(Country, "<br>",
                                               "Men: ", round(LE_men_age,2), "<br>",
                                               "Women: ", round(LE_women_age,2), "<br>",
                                               "Pop: ", format(Total_pop, big.mark = ",")))) +
     geom_jitter(aes(color = Country), alpha = 0.8) +
     labs(title = paste0("Men's vs women's life expectancy in ", sliceYear),
          x = "Life expectancy of men",
          y = "Life expectancy of women") +
     theme_bw() +
     theme(legend.position = "none",
           plot.title = element_text(face = "bold", size = 13))

ggplotly(p2a, tooltip = "text")
```

```{r}
MenVSWomen_regions_yearChosen <- MenVSWomen_regions %>% 
     filter(Year == sliceYear)

p2b <- ggplot(MenVSWomen_regions_yearChosen, aes(x = LE_men_age,
                                                 y = LE_women_age,
                                                 size = Total_pop,
                                                 text = paste0(Region, "<br>",
                                               "Men: ", round(LE_men_age,2), "<br>",
                                               "Women: ", round(LE_women_age,2), "<br>",
                                               "Pop: ", format(Total_pop, big.mark = ",")))) +
     geom_jitter(aes(color = Region), alpha = 0.8) +
     scale_size_continuous(labels = comma) +
     labs(title = paste0("Men's vs women's life expectancy in ", sliceYear),
          x = "Life expectancy of men",
          y = "Life expectancy of women") +
     theme_bw() +
     theme(legend.position = "none",
           plot.title = element_text(face = "bold", size = 13))

ggplotly(p2b, tooltip = "text")
```

```{r}
MenVSWomen_countries_long <- MenVSWomen_countries %>% 
     pivot_longer(cols = c("LE_men_age","LE_women_age"),
                  names_to = "Gender", values_to = "LE_age")
MenVSWomen_countries_long$Gender <- as.factor(sapply(as.character(MenVSWomen_countries_long$Gender), switch,
                                                     "LE_men_age" = "Men",
                                                     "LE_women_age" = "Women"))

# Choose a country
countryChosen <- "United Kingdom"

dfLong_counChosen <- MenVSWomen_countries_long %>% 
     filter(Country == countryChosen) %>% 
     drop_na() %>% 
     as.data.frame()

p3a <- ggplot(dfLong_counChosen, aes(x = Year,
                                     y = ifelse(Gender == "Men",
                                                LE_age, 
                                                -LE_age),
                                     fill = Gender,
                                     text = paste0(Year, "<br>",
                                            "Life exp: ", round(LE_age,2),
                                            " years"))) +
        geom_col() +
        coord_flip() + 
        # scale_fill_manual(labels = c("Men","Women"), values = c("blue","red")) +
        scale_fill_brewer(palette = "Set1") +
        scale_y_continuous(labels = abs,
                           limits = max(dfLong_counChosen$LE_age)*c(-1,1)) +
        labs(title = paste0("Life expectancy by gender in ", countryChosen),
             y = "Life expectancy",
             x = NULL) +
        theme_calc() +
        theme(plot.title = element_text(face = "bold", size = 12))

ggplotly(p3a, tooltip = "text") 
```

```{r}
MenVSWomen_regions_long <- MenVSWomen_regions %>% 
     pivot_longer(cols = c("LE_men_age","LE_women_age"),
                  names_to = "Gender", values_to = "LE_age")
MenVSWomen_regions_long$Gender <- as.factor(sapply(
        as.character(MenVSWomen_regions_long$Gender), switch,
        "LE_men_age" = "Men",
        "LE_women_age" = "Women"))

# Choose a region
regionChosen <- "Africa"

dfLong_regChosen <- MenVSWomen_regions_long %>% 
     filter(Region == regionChosen) %>% 
     drop_na() %>% 
     as.data.frame()

p3b <- ggplot(dfLong_regChosen, aes(x = Year,
                                     y = ifelse(Gender == "Men",
                                                LE_age, 
                                                -LE_age),
                                     fill = Gender,
                                     text = paste0(Year, "<br>",
                                            "Life exp: ", round(LE_age,2),
                                            " years"))) +
        geom_col() +
        coord_flip() + 
        # scale_fill_manual(labels = c("Men","Women"), values = c("blue","red")) +
        scale_fill_brewer(palette = "Set1") +
        scale_y_continuous(labels = abs,
                           limits = max(dfLong_regChosen$LE_age)*c(-1,1)) +
        labs(title = paste0("Life expectancy by gender in ", regionChosen),
             y = "Life expectancy",
             x = NULL) +
        theme_calc() +
        theme(plot.title = element_text(face = "bold", size = 12))

ggplotly(p3b, tooltip = "text") 
```

# Third Dataset

```{r}
medianAge <- as.data.frame(read_csv("./LifeExpectancy/median-age.csv"))
medianAge
```

```{r}
medianAge_countries <- medianAge %>% 
     filter(!is.na(Code) & Entity != "World") %>% 
     rename(Country = Entity, 
            Median_Age = `UN Population Division (Median Age) (2017) (years)`) %>% mutate(Country = as.factor(Country),
            Code = as.factor(Code))
medianAge_regions <- medianAge %>% 
     filter(is.na(Code) | Entity == "World") %>% 
     select(Entity, 
            Year, 
            `UN Population Division (Median Age) (2017) (years)`) %>% 
     rename(Region = Entity, 
            Median_Age = `UN Population Division (Median Age) (2017) (years)`) %>% mutate(Region = as.factor(Region))
medianAge_countries
medianAge_regions
```

5. Menampilkan grafik perbandingan median age antara negara-negara tertentu pada rentang waktu tertentu

```{r}
medianAge_countries_ <- medianAge_countries %>% 
        filter(Country == "China" |
                       Country == "India" |
                       Country == "United States")

ggplot(medianAge_countries_, aes(x = Country,
                                 y = Median_Age)) +
        geom_boxplot() +
        geom_
```

